---
title: "PML_P01_Submission"
author: "Sridhar Somasundharam"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
Loading and preprocessing the data

1. Load the data (i.e. read.csv())
2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r, echo=FALSE}
library(caret)
getwd()
setwd("C:/Users/ssomasundharam/Documents/Data Science/Coursera R/Reproducible Research")
# traincsv<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
# testcsv<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

```

What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day
2. Calculate and report the mean and median total number of steps taken per day

```{r, eval=FALSE}

# download.file(traincsv, destfile = "pml-training.csv")
# download.file(testcsv, destfile = "pml-testing.csv")
# download.file(traincsv, destfile = "pml-training.csv", method = "curl")
# download.file(testcsv, destfile = "pml-testing.csv", method = "curl")

```

```{r}

trainingraw <- read.table("pml-training.csv",sep=",",na.strings = c("NA",""),header=TRUE)
testing <- read.table("pml-testing.csv",sep=",",na.strings = c("NA",""),header=TRUE)

inTrain <- createDataPartition(trainingraw$classe, p=0.70, list=FALSE)
training <- trainingraw[inTrain,]
validation <- trainingraw[-inTrain,]

nearz <- nearZeroVar(training)
training <- training[, -nearz]
validation <- validation[, -nearz]

mostlyNA <- sapply(training, function(x) mean(is.na(x))) > 0.95
training <- training[, mostlyNA==F]
validation <- validation[, mostlyNA==F]


training <- training[, -(1:5)]
validation <- validation[, -(1:5)]

```

### Model building

Fit a model using random forest, running in parallel with 8 processes on i7 the training of the model took ~22 minutes.

```{r, eval=FALSE}

fitControl <- trainControl(method="cv", number=3, verboseIter=F)
fittraining <- train(classe ~ ., data=training, method="rf", trControl=fitControl)
save(fittraining, file="fittraining.RData")

```

```{r, eval=FALSE}

# rm(trainingraw)

load(file = "fittraining.RData")
fittraining$results

```

### Error estimation with cross validation

Using the model that we've trained, we're performing a cross validation with the rest of data from the dataset reserved for this reason. The out of error rate is expected to be less than 1%, as the accuracy of the model observed above is 99.88%.

```{r}

# traincontrol <- trainControl(method = "cv", number = 5)

```

```{r,eval=FALSE}

fitvalidation <- train(classe ~ ., data=validation, method="rf", trControl=fitControl)
save(fitvalidation, file="fitvalidation.RData")

```

```{r,eval=FALSE}

load(file="fitvalidation.RData")
fitvalidation$resample
fitvalidation$results
confusionMatrix(predict(fitvalidation, newdata=validation), validation$classe)

```

Indeed, by calculating the out of sample error (the cross-validation estimate is an out-of-sample estimate) we get the value of `0.54%`:

```{r,eval=FALSE}

fitvalidation$finalModel

```

### Predict the 20 test cases

Finally, to predict the classe of the testing dataset, we're applying the prediction using the model we've trained and output the results in the respective files as adviced by the instructor:

```{r,eval=FALSE}

test_prediction <- predict(fittraining, newdata=testing)
test_prediction
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}
pml_write_files(test_prediction)

```
