---
title: "PML_P01_Submission"
author: "Sridhar Somasundharam"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
Practical Machine Learning is gaining actionable insight, more accurate foresight, and relevant inferences into your data. One of the most common tasks are prediction and machine learning. This project use data which is now possible to collect a large amount of data about personal activity using devices such as Jawbone Up, Nike FuelBand, and Fitbit.  The project will provide basic concepts such as training and tests datasets, over fitting and error rates. The project will also explain the model practical machine learning methods and building prediction based on the training and test datasets. 

Download the data from web site

```{r, echo=FALSE}
library(caret)
getwd()
setwd("C:/Users/ssomasundharam/Documents/Data Science/Coursera R/Practical Machine Learning")
# traincsv<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
# testcsv<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

```

Loading and preprocessing the data

```{r, eval=FALSE}

# download.file(traincsv, destfile = "pml-training.csv")
# download.file(testcsv, destfile = "pml-testing.csv")
# download.file(traincsv, destfile = "pml-training.csv", method = "curl")
# download.file(testcsv, destfile = "pml-testing.csv", method = "curl")

```

Training and test dataset

```{r}

trainingraw <- read.table("pml-training.csv",sep=",",na.strings = c("NA",""),header=TRUE)
testing <- read.table("pml-testing.csv",sep=",",na.strings = c("NA",""),header=TRUE)

inTrain <- createDataPartition(trainingraw$classe, p=0.70, list=FALSE)
training <- trainingraw[inTrain,]
validation <- trainingraw[-inTrain,]

nearz <- nearZeroVar(training)
training <- training[, -nearz]
validation <- validation[, -nearz]

mostlyNA <- sapply(training, function(x) mean(is.na(x))) > 0.95
training <- training[, mostlyNA==F]
validation <- validation[, mostlyNA==F]


training <- training[, -(1:5)]
validation <- validation[, -(1:5)]

```

### Model building

Fit a model using random forest, running in parallel with 8 processes on i7 the training of the model took ~22 minutes.

```{r, eval=FALSE}

fitControl <- trainControl(method="cv", number=3, verboseIter=F)
fittraining <- train(classe ~ ., data=training, method="rf", trControl=fitControl)
save(fittraining, file="fittraining.RData")

```

Save the fit training model

```{r, eval=FALSE}

# rm(trainingraw)

load(file = "fittraining.RData")
fittraining$results

```

### Error estimation with cross validation

Using the model that we've trained, we're performing a cross validation with the rest of data from the dataset reserved for this reason. The out of error rate is expected to be less than 1%, as the accuracy of the model observed above is 99.88%.

```{r}

# traincontrol <- trainControl(method = "cv", number = 5)

```

Save the fit testing model

```{r,eval=FALSE}

fitvalidation <- train(classe ~ ., data=validation, method="rf", trControl=fitControl)
save(fitvalidation, file="fitvalidation.RData")

```

load the fit testing model

```{r,eval=FALSE}

load(file="fitvalidation.RData")
fitvalidation$resample
fitvalidation$results
confusionMatrix(predict(fitvalidation, newdata=validation), validation$classe)

```

Calculating the out of sample error (the cross-validation estimate is an out-of-sample estimate) we get the value of `0.54%`:

```{r,eval=FALSE}

fitvalidation$finalModel

```

### Predict the 20 test cases

Finally, to predict the classe of the testing dataset, we're applying the prediction using the model we've trained and output the results in the respective files as adviced by the instructor:

```{r,eval=FALSE}

test_prediction <- predict(fittraining, newdata=testing)
test_prediction
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}
pml_write_files(test_prediction)

```
